---
title: "infBase Geographic levels"
author: "Clément Turbelin"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{infBase Geographic levels}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Concepts

Theses concepts are used to describe administrative division of a given territory and to provide tools to make operation with these divisions in a generic way.

### Geographic level
  
  A geographic level is a named way to divide a territory in a set of non overlaping area. The full available territory must be divided into area in the level (i.e. the division is complete over the territory).
  
  The concept is abstract from spatial concerns, an "area" of a level can regroup sub-area not spatially connected if needed, as long as the division is consistent (no overlap, and complete). 
  
  A level has a name, used to identify it and the related entities (tables, columns)...
  
### Geographic area  
  
  An area is a element of a level. Identified by a area code (unique in the level), and which can be divided into sub-area in a lower level.
  
### Geographic Hierarchy

A geographic hierarchy path is an ordered list of levels where you can define for each level
 - an upper level (except for the top level)
 - a lower level (except for the bottom level)

For a given area in a level, you can map it to only one area in the upper level and one or more areas in the lower level. In other words a hierarchy should describe a well formed tree with separated branches, with no junction between branches.

## Geographic tables

Each level has a dedicated table defining the list of available areas (territory divisions), called a geographic table.

Geographic tables are named using the pattern 'geo_[name]' where [name] is the level name. For example you have four levels : 'country', 'nut2', 'nuts3', 'zip', you should have a geo_country, geo_nuts2, geo_nuts3, geo_zip.

Each table contains :

- a primary key: id of the area in the level, preferrably named by `code_[name]`
- a title column: name of the area in the level
- ids of the upper levels to allow query accross levels

In order to be consistent you should use names for columns using the name of the corresponding level. Using the example 'code_country', 'code_nuts2', 'code_nuts3', 'code_zip' should be the name of the column for each levels (as primary key for the table of the level and in the other geo_* tables to describe hierarchy).

Following these conventions, the geographic tables should be :

geo_zip      | geo_nut3     | geo_nuts2    | geo_country   |
-----------  |--------------|------------- |-------------  |
code_zip*    | code_nuts3*  | code_nuts2*  | code_country* |
title        | title        | title        | title         |
code_nuts3   | code_nuts2   | code_country |               |
code_nuts2   | code_country |              |               |
code_country |              |              |               | 

Primary keys are marked with the *. It is important that the name of the column refering to a given level must have the exact same name in all geographic tables. The column containing the title should also be the same in every table.

Extract column could be available to describe property of each level.

Platform file should describe available tables and how columns are named. It also handles hierarchies descriptions.

## Configuration

The list of levels & hierarchies should be defined in the *platform file* defining the `geo.levels` object and using the `create_geographic_levels`

For example:
```{r eval=FALSE}
geo.levels = create_geographic_levels(c('zip','nuts3','nuts2','country'), table="geo_zip")
```

Will create a description to a simple geographic structure, with lowest level as "zip", all geographic columns will be be "code_[level]"
and the hierarchy of levels will follow the order given in the list of levels.

The main table used to create cross-levels query will be the lowest level table, named "geo_zip" (describe all area at the "zip" level and had a column for each upper level to map each zip to its upper level).

Of course in this case, "zip" level should be a curated list of zip code, as a zip code can only be related to one area in the upper level (so a choice should has been map to simplify the mapping of zip codes).

geo.tables = describe_geographic_tables(geo.levels)


## Population tables

### Global population
Population tables contains population count for a given level, and a given year

They are named usign the pattern : `pop_[name]`

Table structute should be:
 - code_[name]
 - year
 - population

### Age-group & sex population

Contains by 5-years age group & by sex population for a given year

Table pattern: `pop_age5_[name]`

Table structure:
 - code_[name]
 - year
 - age.min : minimal age of the age group
 - age.max : maximal age of the age group (last group could be NA)
 - all : male + female
 - male: male population
 - female: female population





